---
title: "ag5Tools"  
description: ag5Tools is an R package to download and extract data from the Copernicus \\'Agrometeorological indicators from 1979 to present derived from reanalysis\\' dataset (AgERA5)."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{agera5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## CDSAPI Installation
To download AgERA5 data from the Copernicus Climate Data Store (CDS) you should first follow the steps 
described on the [CDS API Guide](https://cds.climate.copernicus.eu/api-how-to) to install the Python CDSAPI.

## Downloading AgERA5 data
The following example downloads varaible "2m_temperature". For this variable there seven statistics available:  

* 24_hour_maximum

* 24_hour_mean

* 24_hour_minimum

* day_time_maximum

* day_time_mean

* night_time_mean

* night_time_minimum  

In this case, we will download "night_time_minimum". You should replace the target path to 
a location in your own computer where you want to store the downloaded data.

```{r eval = FALSE}
library(ag5Tools)

ag5_download(variable = "2m_temperature",
            statistic = "night_time_minimum",
            day = "all",
            month = "all",
            year = 2015,
            path = "C:/custom_target_folder")

```
